Anthropomorphism in Human-Robot Interaction: Nature, Dynamics and Underlying Cognitive Mechanisms

### 2014-11-25 paper.tex
Polish conclusion and discussion on measuring anthrop.

In this paper, we presented a cognitive and psychological viewpoint on anthropomorphism, what we believe can offer relevant extensions to our understanding of anthropomorphism in HRI. Regarding anthropomorphism as a social phenomenon that arises in an interaction, we formulated the following three suggestions:
\begin{enumerate}
\item \textbf{Anthropomorphism is based on multiple factors.} The three main factors that constitute anthropomorphism, are characteristics of the human $H$, characteristics of the robot $R$, and characteristics of the context of use $C$. A temporal dimension is taken into account by the \textit{interaction history} $t$, which leads to anthropomorphism being also a dynamic phenomenon (see point 3).
\item \textbf{Anthropomorphism is multi-layered.} Applying a cognitive viewpoint on anthropomorphism, there exist \textit{lower and higher levels} of the phenomenon. On one hand, the level (it may be a continuum rather than discrete levels) depends on one how typical or uniquely human the attributed characteristics are, and how much the mental model that the user build of the robot overlaps (or is inspired by) the mental model he/she has of another human. As an outcome of this, three different \textit{cognitive stages} have been identified. These stages reflect the implicated cognitive depth of the anthropomorphic effect.
\item \textbf{Anthropomorphism is dynamic.} Looking at how experience and interaction evolve over time, we developed a conceptual model of the dynamics of anthropomorphism, which states that there is a non-linear relationship between the observable \textit{anthropomorphic effects} $A_t$ (a user's tendency to anthropomorphize a robot) and the \textit{interaction history} $t$ (taking into account the actual time the user has spent interacting with the robot for real or imagined). The shape of the hypothesized curve starts with the so-called \textit{Initial Capital of Anthropomorphism} ($ICA$) that can be measured as a pre-variable, based on the three main factors (see point 1). With a considerable interaction history, the curve approaches the so-called \textit{Stabilized Level of Antrhopomorphism} ($SLA$). How the curve evolves in between these two values is flexible and unique to each user, robot, and context. We generally assume an increase of \textit{anthropomorphic effects} driven by the \textit{novelty} of the robot, followed by a decrease of $A_t$ as the user \textit{familiarizes} him/herself with the robot and with using the robot. 
\end{enumerate}

@@

The Godspeed questionnaire~\citep{bartneck_measurement_2008}, and in particular,
its subpart focused on anthropomorphism, is the main validated questionnaire to
assess anthropomorphism. On 5 point semantic differential scales, people are
asked to rate the following constructs: fake \vs natural, machinelike \vs
humanlike, unconscious \vs conscious, artificial \vs lifelike, moving rigidly
\vs moving elegantly. Because the concept of ``human-likeness'' itself is
complex and abstract, \cite{kahn_jr._robotic_2006} suggest to ask for more
concrete constructs that are typical or unique of the concept of
``human-likeness''.  For instance, \cite{ruijten_introducing_2014} propose a
25-item questions to measure various concrete aspects of human-likeness.

Still, measuring anthropomorphism (or social engagement with a robot) through
subjective self-reporting remains challenging because the phenomenon is
partially unconscious, its mediation is largely subject-dependent, and we found
the phenomenon generally difficult to verbalize for users.

As a consequence, to reduce subjectivity and still accurately account for
psychological effects, it is necessary to interleave quantitative and
qualitative assessments of anthropomorphic effects.

We present in this section the methodology and techniques that we used to assess
anthropomorphic projections in the child-robot interaction study, called the
Domino study (see section \ref{sec:field-results}, and Figure \ref{fig:ranger}).

\paragraph{Qualitative and quantitative assessment of anthropomorphism in
human's behavior\\}

Anthropomorphism can possibly be observed in how humans behave toward a
non-human agent (social human-like engagement). Based on video recordings of the
Domino experiment, we acquired both quantitative and qualitative behavioral data
that reflects anthropomorphism.

Quantitative data are acquired by systematic annotations of specific types of
actions in the video records of the interaction. Six classes of interactions
were identified as reflecting active commitment and/or an anthropomorphic
interaction of the user: gestures toward the robot, touching the robot, direct
talk to the robot, showing something to the robot, mistreating the robot (\eg
kicking it), and looking at the experimenter because worried by the robot's
behavior.

By considering the number of occurrences, the duration, and consequently the
frequency of these types of interaction, we can make statements of how far the
robot is anthropomorphized in an interaction (based on quantitative data).

We assessed anthropomorphic interactions qualitatively by looking at specific
instances of human-social cues in the human's behavior, such as: facial
expressions, observable emotional reactions (\eg surprised, high pitch voice or
shouting to the robot), eye-contact with the robot, turn-taking gestures,
greeting (waving) gestures. These are just suggestions of actions that might
reflect anthropomorphism. These actions partly overlap with the quantitative
assessment, however it is generally advisable to combine the two.

\paragraph{Qualitative assessment of anthropomorphism when verbally referring to
the robot (reflection)\\}

The main tools for qualitative measurement of anthropomorphic effect are
questionnaires and semi-structured interviews. We have explored both
rating-scales in a questionnaire, and open questions in an interview.  In the
Domino study, interviews were transcribed and participant's answers to the key
questions that were assessing an anthropomorphic projection onto the robot, were
analyzed. For instance, one question to see how far children would ascribe moral
standing to the robot was \emph{``If you had this robot at home, and you were to
leave for 2 weeks of holidays, would it be ok to leave the robot alone at
home?''}. A point for anthropomorphism was counted, when the participant's
answer indicated an anthropomorphic view of the robot. This process led to a
``qualitative score'' that reflects how far the robot is anthropomorphized.
Table~\ref{tab:domino-questions} summarizes the constructs that we used to
measure how far the robot is perceived as a human-like agent, along with example
questions.

@@

Some of these questions were taken from previous work done by
\cite{kahn_jr._robotic_2006} an \cite{weiss_i_2009}. In both studies, both
questionnaires and an analysis of behavior were used to assess children's and
adult's engagement with, and perception of the robot. This seems to be a
promising approach to investigate anthropomorphism, as it is also possible to
combine quantitative and qualitative techniques.

@@

Overall, if we agree that \textbf{anthropomorphism arises in an interaction}, this requires studying the phenomenon as such, both in terms of evaluation scenarios, and in terms of measurements. Regarding \textbf{evaluation scenarios}, we suggest to go beyond using videos and pictures of robots, but to use real interaction scenarios (possibly long-term, \eg one interaction spanning over an extended period of time, or several sessions so to have repeated interactions). In terms of \textbf{measurements}, we suggest to go beyond using rating scales, but to apply a more holistic approach that regards not only participants perception of the robot but also the way they interact with the robot. For this, we proposed a so-called \textit{Anthropomorphism Index} $AntX$ that can be used both as a measurement tool and as a value that reflects how much a user anthropomorphized a robot in a given situation. To take into account the dynamics over time, we propose to measure $AntX$ at several points, as a pre-variable (focusing on user expectations, previous experience, robot design), in-situ variables (focusing on behavioral aspects, interaction, variations of robot behavior) and post-variables (perception and acceptance rating of interaction).

@@

We have proposed that the user's familiarization with the robot is a significant factor that influences how anthropomorphism evolves over time. We assume a general \textit{decrease} of anthropomorphic effects as the user gets familiar with the robot, and we applied a cognitive viewpoint to legitimate this. However, there exist opposite interpretations. \cite{eddy_attribution_1993} state that familiarization and more interaction with an agent \textit{increase} the tendency to anthropomorphize. Similarly, \citep{duffy_anthropomorphism_2003} reports that several psychological experiments and HRI studies showed that familiarity with the system eases social acceptance and tends to \textit{increase} people's tendency to perceive human-like characteristics in the system.
We acknowledge that our conceptual model is an idea that needs to be verified, so to see whether our hypotheses can hold. In particular, the effect of the user's familiarity with the robot, and the role of disruptive robot behaviors needs to be studied in more detail and over extended periods of time.

>>>>

In this paper, we have provided a broad account of the cognitive and
psychological viewpoints on anthropomorphism, with a novel focus on the
impact of time. Considering anthropomorphism as a social phenomenon that arises
in an interaction, we formulated the following three suggestions:

\item \textbf{Anthropomorphism is based on multiple factors.} The three main
    factors that constitute anthropomorphism are the characteristics of the
    human, the characteristics of the robot, and the characteristics of the
    context of use. A temporal dimension is taken into account by the
    \textit{interaction history} $t$, which reflects the fact that
    anthropomorphism is also a dynamic phenomenon.

\item \textbf{Anthropomorphism is dynamic.} Looking at how experience and
    interaction evolve over time, we developed a qualitative model of the
    dynamics of anthropomorphism, which materializes a non-linear relationship
    between the observable \textit{anthropomorphic effects} \AntE of the human
    behavior and the \textit{interaction duration} $t$ (the actual time the user
    has spent interacting with the robot for real or imagined). This qualitative
    relationship starts at $t_{0}$ with the \textit{initial potential to
    anthropomorphize} (IPA), a function of the three factors previously
    discussed (point 1) that measures a pre-interaction form of
    anthropomorphism. On the other end, at $t_{\infty}$, we suggest that the
    interaction reaches a \textit{stabilized level of anthropomorphism} (SLA).
    How the curve evolves in between these two values is flexible and unique to
    each user, robot, and context. We however generally hypothesize an increase of
    \textit{anthropomorphic effects} driven by the \textit{novelty} of the
    robot, followed by a decrease of \AntE as the user \textit{familiarizes}
    him/herself with the robot and with using the robot.

\item \textbf{Anthropomorphism is multi-layered cognitive mechanism.} We have
    identified three successive and overlapping cognitive layers that underlie
    human-robot interaction and shape anthropomorphic behaviors.
    These layers (that are likely continuum rather than discrete levels) depend
    on one how typically human the robot is perceived, and how much the mental
    model that the user builds of the robot overlaps (or is inspired by) the
    mental model he/she has of another human. Each of the layers
    reflect different depths of cognitive engagement between the human and the
    robot.

We believe that this systematic approach of anthropomorphism as a complex social
and psychological phenomenon, along with the terminology clarification that we
have proposed in the introduction of the article, can effectively support
further research on the intricate interactions between humans and robots.

We purposefully kept this work at a rather abstract level since our aim was to
build and extend the theoretical grounds of anthropomorphism in HRI. The
question of operationalizing these models is however also important, and the
next section briefly sketch the state of the techniques and methods to
concretely assess anthropomorphic attitudes in HRI.

@@

The Godspeed questionnaire~\citep{bartneck_measurement_2008}, and in particular,
its subpart focused on anthropomorphism, is the main validated questionnaire to
assess anthropomorphism. On 5 point semantic differential scales, people are
asked to rate the following constructs: fake \vs natural, machinelike \vs
humanlike, unconscious \vs conscious, artificial \vs lifelike, moving rigidly
\vs moving elegantly. Because the concept of ``human-likeness'' itself is
complex and abstract, \cite{kahn_jr._robotic_2006} suggest to ask for more
concrete constructs that are typical or unique of the concept of
``human-likeness'', and \cite{ruijten_introducing_2014} propose for instance a
25-item questionnaire to measure various concrete aspects of human-likeness.

We have explored both rating-scales in a questionnaire, and open questions in an
interview.  In the ``dominos'' study depicted in Figure~\ref{fig:ranger}
(page~\pageref{fig:ranger}), interviews with the young children were transcribed
and participant's answers to key questions were analyzed. For instance, one
question to see how far children would ascribe moral standing to the robot was
\emph{``If you had this robot at home, and you were to leave for 2 weeks of
holidays, would it be ok to leave the robot alone at home?''}. We then
attributed points when the participant's answer indicated an anthropomorphic
view of the robot. This process led to a ``qualitative score'' that reflects how
far the robot is anthropomorphized. Table~\ref{tab:domino-questions} summarizes
the constructs that we used to measure how far the robot is perceived as a
human-like agent, along with example questions.

@@

Still, measuring anthropomorphism (or social engagement with a robot) through
subjective self-reporting remains challenging because the phenomenon is
partially unconscious, it remains generally difficult to verbalize for users,
and questionnaires, as post-hoc measurement, do not easily account for the
dynamics of the phenomenon.  Some of the questions of the ``dominos'' study were
taken from previous work done by~\cite{kahn_jr._robotic_2006}
and~\cite{weiss_i_2009}, and in these two studies, questionnaires results were
augmented with an analysis of behavior to assess children's and adult's
engagement with the robot.

We also explored the combination of interviews scoring with behavioral
measurement in the ``dominos'' study: we acquired quantitative data by
systematic annotations of specific types of actions in the video records of the
interaction. Six classes of interactions were identified as reflecting active
commitment and/or an anthropomorphic interaction of the user: gestures toward
the robot, touching the robot, direct talk to the robot, showing something to
the robot, mistreating the robot (\eg kicking it), and looking at the
experimenter because worried by the robot's behavior.
By considering the number of occurrences, the duration, and consequently the
frequency of these types of interaction, we built a weighting scheme to rate
the observed level of anthropomorphic behaviors in an interaction (\ie, \AntE).

This approaches remain however fragile: in our experience, the weighting scheme
of the annotated behaviors mostly relies on a subjective assessment by the
experimenters, which makes it difficult to transfer to different experimental
settings. And it remains to be seen how self-reported assessment of
anthropomorphism (through questionnaires) can be meaningfully combined with
behavioral measurements, in a way that account for the dynamic nature of
anthropomorphism.

@@

Overall, if we agree that \textbf{anthropomorphism arises in an interaction},
this requires studying the phenomenon as such, both in terms of evaluation
scenarios, and in terms of measurements. Regarding \textbf{evaluation
scenarios}, we suggest to go beyond using videos and pictures of robots, but to
use real interaction scenarios (possibly long-term, \eg one interaction spanning
over an extended period of time, or several sessions so to have repeated
interactions). In terms of \textbf{measurements}, we suggest to go beyond using
rating scales, but to apply a more holistic approach that regards not only
participants perception of the robot but also the way they interact with the
robot. To take into account the dynamics over time, it appears important to
assess anthropomorphism at several points, as a
pre-variable (focusing on user expectations, previous experience, robot design),
in-situ variables (focusing on behavioral aspects, interaction, variations of
robot behavior) and post-variables (perception and acceptance rating of
interaction).

@@

We have proposed that the user's familiarization with the robot is a significant
factor that influences how anthropomorphism evolves over time. We assume a
general \textit{decrease} of anthropomorphic effects as the user gets familiar
with the robot, and we applied a cognitive viewpoint to legitimate this.
However, there exist opposite interpretations. \cite{eddy_attribution_1993}
state that familiarization and more interaction with an agent \textit{increase}
the tendency to anthropomorphize. Similarly, \cite{duffy_anthropomorphism_2003}
reports that several psychological experiments and HRI studies showed that
familiarity with the system eases social acceptance and tends to
\textit{increase} people's tendency to perceive human-like characteristics in
the system.  We acknowledge that our conceptual model is an idea that needs to
be verified, so to see whether our hypotheses can hold. In particular, the
effect of the user's familiarity with the robot, and the role of disruptive
robot behaviors needs to be studied in more detail and over extended periods of
time.
